Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
update

\section{Related Work}

@@

\section{Data set}

@@

\section{Methodology}

@@

We chose message and user account features to analyze primarily by the amount of data that we had for each them, while making sure to include the most vital account information. Table \ref{tab:feats} provides averages on a number of the account statistics that we examined. 
\begin{table}
\begin{tabular}{|l|l|r|r|r|}
    \hline
    \textbf{Feature} & \textbf{Label} & \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.}\\
    \hline
    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
    & Ham & 175.51 & 64 & 322.19\\
    \hline
    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
    & Ham & 24.86 & 22 & 10.30\\
    \hline
    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
    & Ham & 15.69 & 6 & 43.40\\
    \hline
    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
    & Ham & 1.89 & 1 & 2.43\\
    \hline
    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
    & Ham & 23.26 & 10 & 58.91\\
    \hline
    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
    & Ham & 25.02 & 22 & 10.70\\
    \hline
    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
    & Ham & 15.01 & 15 & 8.92\\
    \hline
    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
    & Ham & 6.38 & 6 & 3.48\\
    \hline
    \multirow{2}{*}{Profile Desc. Len.} & Spam & 267.30 & 87 & 483.19\\
    & Ham & 439.33 & 267 & 613.26\\
    \hline
    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
    & Ham & 79.29 & 86.00 & 32.85\\
    \hline
\end{tabular}
\caption{Feature statistics}
\label{tab:feats}
\end{table}

@@

@@

In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no explanation for 
this, figure \cite{fig:ysmap} shows that Ghana and Nigeria alone account for 38\% of the IP addresses associated with these 
Yahoo accounts. This popularity of Yahoo among West African users has been noted before\cite{burrell}, and appears to be 
due to the early penetration of Yahoo services in the region.

@@

\section{Classification}

@@

In order to train our classifiers, we had to create a feature matrix from the message and account data. We did this using 
Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
order, mapping each feature (word or expert feature value) to an integer key. The second step was to create a sparse 
matrix representation of each document (in this case, spam or ham message) and the associated features. Spark provides a 
parallelized in-memory computation framework that dramatically reduced the time necessary to extract these features, 
build the dictionary and generate the feature matrix.

@@

%classification:
%- implemented 3 models: naive bayes, linear regression, and SVM.
%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass. The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
%- used scalanlp for naive bayes
%- used matlab for linear regression with ridge regularization.
%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
%
%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
%
%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
%
%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
%

@@

% scalanlp svm

% libsvm

LIBSVM is library for SVM training and classification in C++ and Java, with wrappers for Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters ($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's tools to perform simple scaling on the data.

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is a real-valued function such that: $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

The LIBSVM scaling tool increases file size considerably for data sets with large feature matrices. This large file size, coupled with the computational resources required for SVM training on large data sets, prevented us from completing our analysis of the

\section{Results}

@@

\section{Analysis}
% Vern wants site owner perspective on how to apply classifiers
From a site's perspective, ...

@@

\section{Future Work}

@@

\section{Conclusion}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Related Work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data set}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

@@

We chose message and user account features to analyze primarily by the amount of data that we had for each them, while making sure to include the most vital account information. Table \ref{tab:feats} provides averages on a number of the account statistics that we examined. 
\begin{table}
\begin{tabular}{|l|l|r|r|r|}
    \hline
    \textbf{Feature} & \textbf{Label} & \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.}\\
    \hline
    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
    & Ham & 175.51 & 64 & 322.19\\
    \hline
    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
    & Ham & 24.86 & 22 & 10.30\\
    \hline
    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
    & Ham & 15.69 & 6 & 43.40\\
    \hline
    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
    & Ham & 1.89 & 1 & 2.43\\
    \hline
    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
    & Ham & 23.26 & 10 & 58.91\\
    \hline
    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
    & Ham & 25.02 & 22 & 10.70\\
    \hline
    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
    & Ham & 15.01 & 15 & 8.92\\
    \hline
    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
    & Ham & 6.38 & 6 & 3.48\\
    \hline
    \multirow{2}{*}{Profile Desc. Len.} & Spam & 267.30 & 87 & 483.19\\
    & Ham & 439.33 & 267 & 613.26\\
    \hline
    \multirow{2}{*}{Fields Complete} & Spam & 22.15\% & 14.00\% & 30.75\\
    & Ham & 79.29\% & 86.00\% & 32.85\\
    \hline
    \multirow{2}{*}{Account Lifetime} & Spam & 900.36 h & 1.12 h & 4163.25\\
    & Ham & 8963.19 h & 5297.09 h & 11750.39 \\
    \hline
\end{tabular}
\caption{Feature statistics}
\label{tab:feats}
\end{table}

@@

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/lifetime.pdf}
    \caption{Sender account lifetime at the time that the message was sent.}
    \label{fig:lifetime}
\end{figure}

@@

In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no certain 
explanation for this, Figure \ref{fig:yspam} shows that Ghana and Nigeria alone account for 38\% of the IP 
addresses associated with these Yahoo accounts. This popularity of Yahoo among West African users has 
been noted before\cite{burrell}, and appears to be due to the early penetration of Yahoo services in the region.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Classification}

@@

In order to train our classifiers, we created a feature matrix from the message and account data. We did this using 
Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
order, mapping each feature (word or expert feature value) to an integer key. The second step was to create a sparse 
matrix representation of each document (in this case, spam or ham message) and the associated features. Spark provides a 
parallelized in-memory computation framework that dramatically reduced the time necessary to extract these features, 
build the dictionary and generate the feature matrix.

@@

We initially trained and evaluated three classifiers on a small subset of the data: Na\"ive Bayes, Linear Regression and SVM. 
We chose an 11k spam and 11k ham sample for training, using this particular sample size due to the size of the spam corpus 
section containing recipient age feature data.

%classification:
%- implemented 3 models: naive bayes, linear regression, and SVM.
%- used scalanlp for naive bayes
%- used matlab for linear regression with ridge regularization.
%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
%
%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
%
%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
%
%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
%

@@

We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos algorithm \cite{pegasos} as extended 
by Wang, Crammer and Vucetic\cite{wang2}. The optimizer runs stochastic subgradient descent on the primal objective 
using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
which seems to have led to the classifier's sub-par performance.

We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package'
s tools to perform simple scaling on the data.

Given the complexity and time-intensive nature of choosing and testing parameters, the 
package includes a Python script that performs cross-validation to suggest appropriate 
$\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is 
a real-valued function such that: 
$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis}
% Vern wants site owner perspective on how to apply classifiers
From a site's perspective, ...

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Future Work}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
